Hybrid recurrent model for regime switching nonstationary time series

### 2021-08-30 hybridmodel.tex
Updated title

\lstset{%
	basicstyle={\footnotesize\ttfamily},% footnotesize acceptable for monospace
	numbers=left,numberstyle=\footnotesize,xleftmargin=2em,% show line numbers, remove this entire line if you don't want the numbers.
	aboveskip=0pt,belowskip=0pt,%
	showstringspaces=false,tabsize=2,breaklines=true}
\floatstyle{ruled}
\newfloat{listing}{tb}{lst}{}
\floatname{listing}{Listing}
%
%\nocopyright
%
% PDF Info Is REQUIRED.
% For /Title, write your title in Mixed Case.
% Don't use accents or commands. Retain the parentheses.
% For /Author, add all authors within the parentheses,
% separated by commas. No accents, special characters
% or commands are allowed.
% Keep the /TemplateVersion tag as is
\pdfinfo{
/Title (Hybrid recurrent model for time series with inclusion of external weak signals)
/Author (Etienne David, Jean Bellot, Sylvain Le Corff)
/TemplateVersion (2022.1)
}

@@

% Your title must be in mixed case, not sentence case.
% That means all verbs (including short verbs like be, is, using,and go),
% nouns, adverbs, adjectives should be capitalized, including both words in hyphenated terms, while
% articles, conjunctions, and prepositions are lower case unless they
% directly follow a colon or long dash
\title{Hybrid Recurrent Model With External Weak Signals}
\author{
    %Authors
    % All authors must be in the same font size and format.
    Etienne David,\textsuperscript{\rm 1,\rm 2}\\
    Jean Bellot,\textsuperscript{\rm 1}\\
    Sylvain Le Corff\textsuperscript{\rm 2}\\
}
\affiliations{
    %Afiliations
    \textsuperscript{\rm 1}Heuritech, Paris\\
    \textsuperscript{\rm 2}Samovar, T\'el\'ecom SudParis, d\'epartement CITI, TIPIC, Institut Polytechnique de Paris, Palaiseau
    % If you have multiple authors and multiple affiliations
    % use superscripts in text and roman font to identify them.
    % For example,

@@

\begin{abstract}
Developing models and algorithms to draw causal inference for time series is a long standing statistical problem. It is crucial for many applications, in particular for fashion or retail industries, to make optimal inventory decision and avoid massive wastes. By tracking thousands of fashion trends on social media with stat-of-the-art computer vision approaches, we propose in this paper, the first study of fashion time series forecasting. Our contribution is  twofold. We first provide the first fashion dataset gathering \numberts\ weekly fashion time series. As influence dynamics are the key of emerging trend detection, we associate with each time series several external weak signals representing behaviors of influencers towards the trend. Secondly, to leverage such a complex and rich dataset, we propose a new hybrid forecasting model. Our approach combines per-time-series parametric models with seasonal components and a global recurrent neural network to include sporadic external signals. This hybrid model provides state-of-the-art results on the fashion dataset, and illustrates the benefit of the contribution of  external weak signals.
\end{abstract}

@@

\subsection{Per time series statistical model}
%The first step in our additive hybrid model is to consider, for each time series, an exponential smoothing  statistical model.
The time-series-specific models compute, for each sequence, two components,  level  and  seasonality. For all $1\leqslant n \leqslant N$ and $0\leqslant t \leqslant T$, let $\lvl_t^n$ be the value of the $n$-th level component at time $t$ and  $\fulllvl^n = \{\lvl_t^n\}_{0\leqslant t \leqslant T}$ be all the values of this level component. For all $1\leqslant n \leqslant N$ and $-m+1\leqslant t \leqslant T$, let $\seas_t^n$ be the value of the $n$-th seasonality component at time $t$ and  $\fullseas^n = \{\seas_t^n\}_{-m+1\leqslant t \leqslant T}$ be all the values of this season component. $m \in \mathbb{N}$ is the seasonality (4 for quarterly, 12 for monthly or 52 for weekly time series). Given the sequence $\fullts^n$, the two components  $\fulllvl^n$ and $\fullseas^n$ are computed using exponential smoothing.
For all $1\leqslant n \leqslant N$, let $\Theta^n = \{\alpha^n, \gamma^n, \lvl^n_0, \{\seas_t^n\}_{-m+1\leqslant t \leqslant 0}\}$ be the parameters of the $n$-th sequence with $\lsmooth^n$ and $\ssmooth^n$ the positive smoothing parameters, $\lvl^n_0$ the first level term and $\{\seas_t^n\}_{-m+1\leqslant t \leqslant 0}$ the $m$ first seasonality terms. We have finally for all $1\leqslant n \leqslant N$ and $1\leqslant t \leqslant T$, \textcolor{blue}{Now, I initialize alpha, gamma, the 52 first seasons and the first level with a python package}
%We initialize the first level with $\lvl^n_{0} = \sum_{t = 1}^m \ts^n_{t}/m$ \textcolor{red}{issue with the time index ?} and, for all $1\leqslant n \leqslant N$ and $1\leqslant t \leqslant T$,
\begin{align*}
    \lvl^n_{t} &= \lsmooth^n (\ts^n_t - \seas^n_{t-m}) + (1- \lsmooth^n)\lvl^n_{t-1}\,, \\
    \seas^n_t &= \ssmooth^n (\ts^n_t - \lvl^n_{t-1}) + (1 - \ssmooth^n)\seas^n_{t-1}\,. 
\end{align*}

For all $1\leqslant n \leqslant N$, given the level component $\fulllvl^n$, the seasonality component $\fullseas^n$ and the forecast horizon $\lag \in \mathbb{N}$, the exponential smoothing forecast $\{\etspred^{ets,n}_{T+1:T+\lag|T}\}_{1\leqslant n \leqslant N}$ is, for all $i \in \{1,\ldots,\lag\}$,
$$
\etspred^{ets,n}_{T+i|T} = \lvl^n_T + \seas^n_{T+i-m(k+1)}\,,
$$
where $k = \lfloor(i-1)/m\rfloor$.

>>>>

\lstset{%
	basicstyle={\footnotesize\ttfamily},% footnotesize acceptable for monospace
	numbers=left,numberstyle=\footnotesize,xleftmargin=2em,% show line numbers, remove this entire line if you don't want the numbers.
	aboveskip=0pt,belowskip=0pt,%
	showstringspaces=false,tabsize=2,breaklines=true}
\floatstyle{ruled}
\newfloat{listing}{tb}{lst}{}
\floatname{listing}{Listing}
%
%\nocopyright
%
% PDF Info Is REQUIRED.
% For /Title, write your title in Mixed Case.
% Don't use accents or commands. Retain the parentheses.
% For /Author, add all authors within the parentheses,
% separated by commas. No accents, special characters
% or commands are allowed.
% Keep the /TemplateVersion tag as is
\pdfinfo{
/Title (Hybrid recurrent model for regime switching nonstationary time series)
/Author (Etienne David, Jean Bellot, Sylvain Le Corff)
/TemplateVersion (2022.1)
}

@@

% Your title must be in mixed case, not sentence case.
% That means all verbs (including short verbs like be, is, using,and go),
% nouns, adverbs, adjectives should be capitalized, including both words in hyphenated terms, while
% articles, conjunctions, and prepositions are lower case unless they
% directly follow a colon or long dash
\title{Hybrid recurrent model for regime switching nonstationary time series}
\author{
    %Authors
    % All authors must be in the same font size and format.
    Etienne David,\textsuperscript{\rm 1,\rm 2}\\
    Jean Bellot,\textsuperscript{\rm 1}\\
    Sylvain Le Corff\textsuperscript{\rm 2}\\
}
\affiliations{
    %Afiliations
    \textsuperscript{\rm 1}Heuritech, Paris\\
    \textsuperscript{\rm 2}Samovar, T\'el\'ecom SudParis, d\'epartement CITI, TIPIC, Institut Polytechnique de Paris, Palaiseau
    % If you have multiple authors and multiple affiliations
    % use superscripts in text and roman font to identify them.
    % For example,

@@

\begin{abstract}
Developing models and algorithms to draw causal inference for time series is a long standing statistical problem. It is crucial for many applications, in particular for fashion or retail industries, to make optimal inventory decision and avoid massive wastes. By tracking thousands of fashion trends on social media with stat-of-the-art computer vision approaches, we propose a new model for fashion time series forecasting. Our contribution is  twofold. We first provide the first fashion dataset gathering \numberts\ weekly fashion time series. As influence dynamics are the key of emerging trend detection, we associate with each time series several external weak signals representing behaviors of influencers towards the trend. Secondly, to leverage such a complex and rich dataset, we propose a new hybrid forecasting model. Our approach combines per-time-series parametric models with seasonal components and a global recurrent neural network to include sporadic external signals. This hybrid model provides state-of-the-art results on the fashion dataset, and illustrates the benefit of the contribution of  external weak signals.
\end{abstract}

@@

\subsection{Per time series statistical model}
%The first step in our additive hybrid model is to consider, for each time series, an exponential smoothing  statistical model.
The time-series-specific models compute, for each sequence, two components,  level  and  seasonality. For all $1\leqslant n \leqslant N$ and $0\leqslant t \leqslant T$, let $\lvl_t^n$ be the value of the $n$-th level component at time $t$ and  $\fulllvl^n = \{\lvl_t^n\}_{0\leqslant t \leqslant T}$ be all the values of this level component. For all $1\leqslant n \leqslant N$ and $-m+1\leqslant t \leqslant T$, let $\seas_t^n$ be the value of the $n$-th seasonality component at time $t$ and  $\fullseas^n = \{\seas_t^n\}_{-m+1\leqslant t \leqslant T}$ be all the values of this season component. $m \in \mathbb{N}$ is the seasonality (4 for quarterly, 12 for monthly or 52 for weekly time series). Given the sequence $\fullts^n$, the two components  $\fulllvl^n$ and $\fullseas^n$ are computed using exponential smoothing.
For all $1\leqslant n \leqslant N$, let $\Theta^n = \{\alpha^n, \gamma^n,\{\seas_t^n\}_{-m+1\leqslant t \leqslant 0}\}$ be the parameters of the $n$-th sequence with $\lsmooth^n$ and $\ssmooth^n$ the positive smoothing parameters and $\{\seas_t^n\}_{-m+1\leqslant t \leqslant 0}$ the $m$ first seasonality terms. We initialize the first level with $\lvl^n_{0} = \sum_{t = 1}^m \ts^n_{t}/m$ \textcolor{red}{issue with the time index ?} and, for all $1\leqslant n \leqslant N$ and $1\leqslant t \leqslant T$,
\begin{align*}
    \lvl^n_{t} &= \lsmooth^n (\ts^n_t - \seas^n_{t-m}) + (1- \lsmooth^n)\lvl^n_{t-1}\,, \\
    \seas^n_t &= \ssmooth^n (\ts^n_t - \lvl^n_{t-1}) + (1 - \ssmooth^n)\seas^n_{t-1}\,. 
\end{align*}

For all $n \in \{1,\ldots,N\}$, given the level component $\fulllvl^n$, the seasonality component $\fullseas^n$ and the forecast horizon $\lag \in \mathbb{N}$, the exponential smoothing forecast $\{\etspred^{ets,n}_{T+1:T+\lag|T}\}_{1\leqslant n \leqslant N}$ is, for all $i \in \{1,\ldots,\lag\}$,
$$
\etspred^{ets,n}_{T+i|T} = \lvl^n_T + \seas^n_{T+i-m(k+1)}\,,
$$
where $k = \lfloor(i-1)/m\rfloor$.
